---
title: "recapr_exercise"
author: "Kyle Burt"
format: html
editor: visual
---

# Libraries

```{r}
library(here)
library(tidyverse)
library(recapr)
```

'recapr' package https://cran.r-project.org/web/packages/recapr/recapr.pdf

# EXAMPLE

```{r}
n1<-75 # number of individuals marked in the first sampling
n2<-100 # number of individuals captured in the second sampling effort
m2<- 4 # number of individuals in the second sampling effort that were marked.
```

### Runs the model, get the estimate of N and calculates a confidence interval

```{r}
NPetersen(n1, n2, m2) # estimates the population size
ciPetersen(n1,n2,m2)  # 95% CIs using normal approximation ($ciNorm) and bootstrapping ($ciBoot)
```

# 1. Run the Petersen Estimator for the Silver Carp parameters

### Example from lecture notes; Sass et al. (2010) Silver carp study

```{r}
n1_sc <- 4540
n2_sc <- 2239 #approximated from the paper (not reported)
m2_sc <- 30
```

### Chapman estimator

```{r}
NChapman(n1_sc, n2_sc, m2_sc)
ciChapman(n1_sc, n2_sc, m2_sc)
```

### Petersen estimator

```{r}
NPetersen(n1_sc, n2_sc, m2_sc) 
ciPetersen(n1_sc, n2_sc,m2_sc)
```

### 2. What is the difference between the Petersen and Chapman estimates and the confidence intervals

-   Difference in Estimates: 338,835.3 - 328,122.9 = 10,712.4 individuals

-   Petersen confidence intervals: Normal (218,401.9 to 459,268.7); Bootstrap (247,928.3 to 508,253.0)

-   Chapman confidence intervals: Normal (215,611 to 440,635); Bootstrap (242,186 to 484,372)

# PORTFOLIO ASSIGNMENT

N1 = 78 N2 = 168 m2 = 19

```{r}
n1_mm <- 78
n2_mm <- 168
m2_mm <- 19
```

```{r}
NChapman(n1_mm, n2_mm, m2_mm)
ciChapman(n1_mm, n2_mm, m2_mm)
```

```{r}
NPetersen(n1_mm, n2_mm, m2_mm) 
ciPetersen(n1_mm, n2_mm,m2_mm)
```

### 3. How many M&Ms are in this bag?

-   Chapman estimate = 666.55

-   Petersen estimate = 689.6842
